---
title: "Investigating the Relationship between Physical Activity and an Individual's Study Habits"
author: "Simon Kim"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary

This report aims to investigate how physical activity impacts study habits.

Key findings:

- The more time spent performing physical activity, the longer respondents were able to undergo focused study.

- The way individuals perform physical activity has strong impacts on how energised they were before study.


<br>

# Full Report

## Initial Data Analysis (IDA)

### Source

Data was collected using (Google Forms)[<https://docs.google.com/forms/d/e/1FAIpQLSdoKyCJjY2eqYijesKuMEWtXnKf2gN7jZkD40PZ7peMhT8kPw/viewform>]

```{r, message=F}

data <- read.csv("data.csv")

# loading packages

library(dplyr)
library(plotly)

# show unclean data
str(data)

```


### Data Cleaning

```{r}
# incorrect data
data$X <- NULL

# rename variables for ease of use
data <- rename(data, "physical_activity_status" = "X1..Do.you.partake.in.physical.activity.")
data <- rename(data, "physical_activity_days_per_week" = "X2..How.many.days.per.week.do.you.partake.in.physical.activity.")
data <- rename(data, "physical_activity_hours_per_week" = "X3..How.many.hours.of.physical.activity.do.you.partake.in.each.week.")
data <- rename(data, "intensity" = "X4..How.would.you.describe.the.intensity.of.your.physical.activity.")
data <- rename(data, "type" = "X5..What.type.of.physical.activity.do.you.partake.in.")
data <- rename(data, "study_before_after" = "X6..Do.you.tend.to.study.before.or.after.physical.activity.")
data <- rename(data, "energised_refreshed_status" = "X7..Do.you.generally.feel.energised.and.refreshed.before.study.")
data <- rename(data, "study_hours_per_week" = "X8..How.many.hours.of.study.do.you.undergo.per.week.")
data <- rename(data, "focused_study" = "X9..How.many.hours.of.focused.study.can.you.undergo.in.one.sitting.")

# chr -> factor for ease of use
data$physical_activity_status <- as.factor(data$physical_activity_status)
data$intensity <- as.factor(data$intensity)
data$type <- as.factor(data$type)
data$study_before_after <- as.factor(data$study_before_after)
data$energised_refreshed_status <- as.factor(data$energised_refreshed_status)
data$focused_study <- as.factor(data$focused_study)

# show cleaned data
str(data)

```


### Structure

Using the survey, qualitative and quantitative data was collected through the use of multiple-choice and short-answer questions. There are 9 variables and 37 survey responses.


### Limitations

-   Selection bias: The majority of respondents are students at USYD, and many show different survey responses to those who are not university students in USYD.

-   Consent Bias: The mode of data collection (survey) inherently introduces consent bias because it gives respondents the choice on whether or not to participate.


### Assumptions

-   This study assumes that respondents partake in some amount of study.


<br>


## Research Question(s)

### Research Question 1

> Does the amount of time spent training influence the amount and quality of study?

```{r, message=F}

# physical_activity_days_per_week vs frequency barplot
plot1 <- ggplot(data=data,
                aes(x=physical_activity_days_per_week)) +
          geom_bar() +
          xlab("Number of days per week physical activity is done") +
          ylab("Frequency") +
          theme_bw() +
          scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7))


# physical_activity_hours_per_week vs study_hours_per_week on scatter 
plot2 <- ggplot(data=data,
                  aes(x=physical_activity_hours_per_week,
                      y=study_hours_per_week)) +
            geom_point() +
            xlab("Number of hours per week physical activity is done") +
            ylab("Number of hours per week study is done") +
            theme_bw() +
            geom_smooth(method="lm")

cor(data$physical_activity_hours_per_week, data$study_hours_per_week)
          

# physical_activity_days_per_week vs study_hours_per_week on scatter
plot3 <- ggplot(data=data,
                  aes(x=physical_activity_days_per_week,
                      y=study_hours_per_week)) +
            geom_point() +
            xlab("Number of days per week physical activity is done") +
            ylab("Number of hours per week study is done") +
            theme_bw() +
            geom_smooth(method="lm")

cor(data$physical_activity_days_per_week, data$study_hours_per_week)


# physical_activity_days_per_week vs focused_study barplot
plot4 <- ggplot(data=data,
                aes(x=physical_activity_days_per_week,
                    fill=focused_study)) +
          geom_bar() +
          xlab("Number of days per week physical activity is done") +
          ylab("Frequency") +
          theme_bw() +
          scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7))


# focused_study vs physical_activity_hours_per_week barplot
plot5 <- ggplot(data=data,
                  aes(x=focused_study,
                      y=physical_activity_hours_per_week)) +
            geom_boxplot() +
            xlab("Number of hours of focused study completed in one sitting") +
            ylab("Number of hours per week physical activity is done") +
            theme_bw() +
            scale_x_discrete(limits=c("< 1 hours", "1-2 hours", "2-3 hours", "3-4 hours", "> 4 hours"))



plotly::ggplotly(plot1)
plotly::ggplotly(plot2)
plotly::ggplotly(plot3)
plotly::ggplotly(plot4)
plotly::ggplotly(plot5)

```

<br>

#### Analysis and Summary

- Respondents who performed physical activity a greater number of hours per week spent fewer hours per week studying (plot2). This trend was reflected when observing how many days per week physical activity was performed (plot3). A possible explanation for this could be the time commitment of physical activity, but it is difficult to be certain without further investigation.

- However, respondents who performed physical activity a greater number of days per week were able to undergo longer periods of focused study in one sitting compared to respondents who performed physical activity a fewer number of days per week (plot4). This trend was reflected when observing how many hours per week of physical activity were performed (plot5). Research shows a strong correlation between physical activity and concentration.

- Overall, there was an inverse effect when observing how the number of days per week physical activity is performed impacts; the number of hours per week study, and the number of hours of focused study in one sitting.

<br>

#### Linear Modelling

```{r, message=F}

model <- lm(data$study_hours_per_week ~ data$physical_activity_hours_per_week)

summary(model)

```

```{r, message=F}

residual_plot <- ggplot(data=model,
                        aes(x=.fitted,
                            y=.resid)) +
                  geom_point() +
                  geom_hline(yintercept=0, linetype="dashed", colour="red") +
                  ggtitle("Residual Plot") +
                  xlab("Time spent studying per week (hours)") +
                  ylab("Residuals") +
                  theme_bw()


plotly::ggplotly(residual_plot)

```

##### Analysis

- Due to the heteroscedasticity of the residual plot, it would not be appropriate to observe and analyse it.

<br>

### Research Question 2

> How does the way an individual trains influence how refreshed they are before study?

#### Type of Physical Activity

```{r}

# type barplot with evergised_refreshed_status fill
plot6 <- ggplot(data=data,
                aes(x=type,
                    fill=energised_refreshed_status)) +
          geom_bar() +
          xlab("Type of physical Activity") +
          ylab("Frequency") +
          theme_bw()


plotly::ggplotly(plot6) # more useful plot 

```


##### Analysis

- Respondents who performed physical activity of the types: resistance training or flexibility training were more energetic before study. 

- Respondents who performed physical activity of the type: incidental activity were less energetic before study.

- Respondents who performed physical activity of the type: cardiovascular training showed minimal impact on how energetic they were before study.

<br>

#### Intensity of Physical Activity

```{r}

# intensity barplot with energised_refreshed_status fill
plot7 <- ggplot(data=data,
                  aes(x=intensity,
                      fill=energised_refreshed_status)) +
            geom_bar() +
            xlab("Intensity of physical activity") +
            ylab("Frequeny") +
            theme_bw() +
            scale_x_discrete(limits=c("Low intensity", "Moderate intensity", "High intensity", "High performance athlete intensity"))


plotly::ggplotly(plot7)

```


##### Analysis and Summary

- The higher the intensity of physical activity, the more energetic respondents were before study.

- A limitation is that there is less data on respondents who performed physical activity at a high performance athlete intensity - this can be partially explained due to the fact that there are less people who train at such a high level compared to people who train at a lower intensity. Thus, it is difficult to provide  trends when observing respondents who perform at a high performance athlete intensity.

- Overall, respondents who performed physical activity at higher intensities showed higher levels of energy before study, but further investigation is required.

<br>

#### Study Before or After Physical Activity

```{r, message=F}

# study_before_after barplot with energised_refreshed_status fill
plot8 <- ggplot(data=data,
                aes(x=study_before_after,
                    fill=energised_refreshed_status)) +
          geom_bar() +
          xlab("Study before or after physical activtiy") +
          ylab("Frequency") +
          theme_bw()


plotly::ggplotly(plot8)

```


##### Analysis and Summary

- Studying before or after physical activity has little impact on how refreshed and energetic respondents were before studying.
